---
title: "Data Analyses"
subtitle: "Palaeontological evidence shows long-term declines in Red Sea coral colony size"
author: 
  - Lewis A. Jones
  - Angelina Ivkić
  - Andreas Kroh
  - Abbas Mansour
  - Mohamed Osman
  - Mohamed Hassan
  - Martin Zuschin
institute: 
  - Centro de Investigación Mariña, Departamento de Ecoloxía e Bioloxía Animal, Universidade de Vigo, 36310 Vigo, Spain
  - Department of Palaeontology, University of Vienna, Josef-Holaubek-Platz 2, 1090 Vienna, Austria
  - Natural History Museum Vienna, Department of Geology & Palaeontology, 1010, Vienna, Austria
  - South Valley University, Department of Geology, Qena, Egypt
  - South Valley University, Department of Geology, Qena, Egypt
  - South Valley University, Department of Geology, Qena, Egypt
  - Department of Palaeontology, University of Vienna, Josef-Holaubek-Platz 2, 1090 Vienna, Austria
date: today
toc: true
format:
  html:
    html-math-method: katex
    code-tools: true
    self-contained: true
knitr:
  opts_chunk:
    out.width: "100%"
    out.height: "100%"
    fig.path: ../article/figures/
execute:
  warning: false
---

## Setup

Load relevant packages and set default options.

```{r, setup, include=TRUE}
# Package for data handling/manipulation
library(tidyverse)
# Package for colour palettes
library(MetBrewer)
# Package for generic plotting
library(ggplot2)
# Package for adding scale bar and north arrow
library(ggspatial)
# Package for text labels
library(ggrepel)
# Package for combining plots
library(cowplot)
# Packages for getting map data
library(rnaturalearth)
library(rnaturalearthdata)
# Set root directory
knitr::opts_knit$set(root.dir = '..')
```

## Study sites

Generate a map of modern and Last Interglacial fossil sites.



```{r, map}
#| fig-height: 7
# Read modern data
modern <- read.csv("data-raw/corals_egypt.csv")
# Read fossil data
fossil <- read.csv("data-raw/corals_egypt.csv")
# Get unique sites
modern_sites <- unique(modern[, c("lng", "lat", "site")])
modern_sites$age <- "Modern"
fossil_sites <- unique(fossil[, c("lng", "lat", "site")])
fossil_sites$age <- "Last Interglacial"
# Add ages
sites <- rbind.data.frame(modern_sites, fossil_sites)
# Define factor levels
sites$age <- factor(sites$age, levels = c("Modern", "Last Interglacial"))
# World sf
world <- ne_countries(scale = "large", continent = "Africa",
                      returnclass = "sf")
# Define countries for plotting
countries <- c("egypt", "saudi arabia", "jordan", "isreal", "syria",
               "lebanon", "iraq", "iran", "kuwait")
# Get country sfs
countries <- ne_countries(scale = "large", country = countries,
                          returnclass = "sf")

# Define colours for plotting
land <- "darkgrey"
sea <- "aliceblue"

# Make map of Africa
africa <- ggplot(data = world) +
  geom_sf(colour = land, fill = land) +
  geom_sf(data = countries, colour = land, fill = land) +
  geom_rect(aes(xmin = 33, xmax = 38, ymin = 24, ymax = 28),
            fill = NA, colour = "black", linewidth = 0.3) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  coord_sf(xlim = c(-19, 50)) +
  theme_bw() +
  theme(panel.background = element_rect(fill = sea),
        plot.background = element_blank(),
        axis.ticks = element_blank(),
        axis.text = element_blank(),
        axis.title = element_blank(),
        panel.grid = element_blank())

# Make locality map
localities <- ggplot(data = countries) +
  geom_sf(colour = "black", fill = land) +
  geom_point(data = sites, aes(x = lng, y = lat, fill = age),
             shape = 21, size = 4,
             colour = "black", alpha = 0.75) +
  geom_text_repel(data = sites, aes(x = lng, y = lat, label = site,
                                    colour = age),
                  size = 3, min.segment.length = unit(0, 'cm'),
                  box.padding = 1,
                  nudge_x = 0.5) +
  scale_x_continuous(expand = c(0, 0)) +
  scale_y_continuous(expand = c(0, 0)) +
  # Add colour scale
  scale_colour_met_d("Hiroshige") +
  # Add fill scale
  scale_fill_met_d("Hiroshige") +
  xlab("Longitude (º)") +
  ylab("Latitude (º)") +
  annotation_scale(text_cex = 1) +
  annotation_north_arrow(location = "bl",
                         pad_y = unit(1, "cm"),
                         height = unit(2, "cm"), width = unit(2, "cm")) +
  coord_sf(xlim = c(33, 38), ylim = c(24, 28)) +
  theme_bw() +
  theme(panel.background = element_rect(fill = sea),
        plot.background = element_blank(),
        axis.title = element_text(size = 16),
        axis.text = element_text(size = 12),
        panel.grid = element_blank(),
        legend.position = "top",
        legend.title = element_blank())

# Create inset
map <- ggdraw(localities) +
  draw_plot(
    {
      africa
    },
    x = 0.748, y = 0.6385,
    width = 0.3, height = 0.25
  )
map
```

Load Last Interglacial transect data and filter to intercepts identified to genus or species level. Subsequently, calculate the median intercept and 5^th^ and 95^th^ percentiles for each taxa, as well as the number of intercepts. Finally, filter the dataset to retain only taxa with 5 or more intercepts.

```{r, prepare-data}
# Plot density distributions
fossil %<>% 
  # Filter to Scleractinia
  filter(order == "Scleractinia") %>%
  # Retain intercepts IDed to genus/species-level
  filter(rank == "genus" | rank == "species") %>%
  # Retain genus and intercept length
  select("genus", "length") %>% 
  # Group by genus
  group_by(genus) %>% 
  # Calculate median per genus
  mutate(lower = quantile(length, probs = 0.05),
         average = median(length),
         upper = quantile(length, probs = 0.95)) %>%
  # Count number of intercepts per genus
  add_count() %>% 
  # Add age
  add_column(age = "Last Interglacial") %>%
  # Retain taxa with 5 or more intercepts
  filter(n >= 5)
```

Generate a density plot of the distribution of coral colony sizes for each genus and time.

```{r, taxonomic-plot}
#| fig-height: 7
# Plot data
ggplot(data = fossil, aes(x = length, y = ..count..)) +
  # Plot density
  geom_density(stat = "density", aes(fill = age, colour = age), alpha = 0.6) +
  # Add vertical line of median value
  geom_vline(aes(xintercept = average, colour = age), linetype = 2) +
  # Add points of the median value
  geom_point(aes(x = average, y = n / 2, fill = age), colour = "black", shape = 21) +
  # Transform x-axis to log10
  scale_x_continuous(trans = "log10", limits = c(1, 600)) +
  # Add colour scale
  scale_colour_met_d("Hiroshige") +
  # Add fill scale
  scale_fill_met_d("Hiroshige") +
  # Y-axis lavel
  ylab(lab = "number of intercepts") +
  # X-axis label
  xlab(lab = "colony size (cm)") +
  # Create facets across taxa with free scales
  facet_wrap(~genus, ncol = 4, scales = "free") +
  # Set themes
  theme_bw() +
  theme(
    legend.title = element_blank(),
    legend.position = "top"
  )
```


# Number of intercepts per species

# Coral colony size frequency distributions
Visualise size distributions or cumulative frequency plots 
Kolmogorov–Simirnov (KS) comparison of distributions.

# Box-plot of size distributions between genera

# Box plot between reef edge and reef slope modern/LIG


Geometric mean... log data
